Quantifying Limits to Detection of Early Warning for Critical Transitions

### 2012-01-23 boettiger.tex
minor edits

\section{A model based approach}
Model-based approaches are beginning to play a larger role in early warning signals, 
though we have not as yet seen the direct fitting and simulation of models to compare hypotheses.  
While choosing appropriate models without system-specific knowledge is challenging, 
much can be accomplished by framing the implicit assumptions above into more precise equations.
\citet{Lade2011} introduces the idea of generalized models for early warning signals, and 
\citet{Kuehn2011} presents normal forms for bifurcation processes that can give rise to critical transitions.  
While presented as a ``non-parametric'' approach,~\citet{Carpenter2011e} is essentially similar, where the dynamics are approximated 
by an explicit stochastic differential equation (SDE) whose coefficients are estimated from statistical moments of the data.

\cb{We probably need to give some sense where this section is headed, so here's my minimal summary}

We introduce a set of SDE models that is closely related to these examples,
and go on to illustrate how these models can be fit to data by maximum likelihood
and compared by a simulation or bootstrapping approach rooted in~\citet{Cox1961} and~\citet{McLachlan1987} that
characterizes the rate of missed detections and false alarms expected in the estimate. 
\cb{or equivalently, could say: estimates both the power and statistical significance of the test.}  
\cb{Is it clear that the maximum likelihood fitting and the simulation/bootstrapping are novel here, as is the power? }

>>>>

\section{A model based approach}
Model-based approaches are beginning to play a larger role in early warning signals, 
though we have not as yet seen the direct fitting and simulation of models to compare hypotheses.  
While choosing appropriate models without system-specific knowledge is challenging, 
much can be accomplished by framing the implicit assumptions above into more precise equations.
\citet{Lade2011} introduces the idea of generalized models for early warning signals, and 
\citet{Kuehn2011} presents normal forms for bifurcation processes that can give rise to critical transitions.  
While presented as a ``non-parametric'' approach,~\citet{Carpenter2011e} is essentially similar, where the dynamics are approximated 
by an explicit stochastic differential equation (SDE) whose coefficients are estimated from statistical moments of the data. \citet{Dakos2011a} starts by explicitly assuming these same SDE models but uses them only to define the summary statistics.  
% If we are to build are methods on the assumptions

\cb{We probably need to give some sense where this section is headed, so here's my minimal summary}
We introduce a set of SDE models that is closely related to these examples,
and go on to illustrate how these models can be fit to data by maximum likelihood
and compared by a simulation or bootstrapping approach rooted in~\citet{Cox1961} and~\citet{McLachlan1987} that
characterizes the rate of missed detections and false alarms expected in the estimate. 
\cb{or equivalently, could say: estimates both the power and statistical significance of the test.}  
\cb{Is it clear that the maximum likelihood fitting and the simulation/bootstrapping are novel here, as is the power? }
